---
title: "sRm"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{sRm}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

### Introduction

`sRm` is package which contains a selection of functions and methods for handling, processing and visualising Selective Reaction Monitoring (SRM) or Multiple Reaction Monitoring (MRM) Liquid Chromatography Mass Spectrometry Data (LC-MS). The key distinction between SRM/MRM data and other LC-MS data is that the data is acquired as a time-series in the form of chromatograms; whereas other LC-MS (ie DIA, DDA, PRM) acquire data as a series of mass specturms which can then be used to create chromatograms. 

`sRm` is designed to make use of existing packages (ie, much of the heavy lifting is done by `mzR` and `xcms`) and create a simple S4 interface to the user for processing, visualisation and reporting. 


**`sRm` can be easily installed directly from GitHub using the `remotes` package.**

```{r, install, eval=FALSE}
remotes::install_github('wilsontom/sRm')
```

### Loading Data

Open and parse a series of `.mzML` files using the `openSRM` function. 

```{r, load}
library(sRm)

chrom_files <- list.files(system.file('extdata/Thermo_RAW', package = 'sRm'), full.names = TRUE)

srmExp <- openSRM(chrom_files, source_type = 'raw')

srmExp

```

The `SRM` object contains the following data;

 * **meta**
    * a `tibble` of a minimal amount of meta data.
 * **transitions**
    * a `tibble` of all transitions names. Transitions are all given a numeric index to make accessing transitions easier for plotting.
 * **rawChrom**
    * a `tibble` of retention time (rt) and intensity (int) values for all imported SRM chromatograms.
 * **transformedChrom**
 * **peaks**
    * a `tibble` of detected peaks for each SRM chromatogram.
 * **header**
    * a `tibble` of formated header data from `mzR::header()`


### Plotting Data

#### plotSRM

The main feature of `sRm` is a series of plotting methods which are useful for visualising the raw data. The `plotSRM` method uses the numeric index of the transition. These can be found in the `transitions` slot. The `plotSRM` method has two `type` options; overlay and facet. 

```{r, plotSRM_overlay, fig.width = 6, fig.height = 4, fig.align = 'center'}
plotSRM(srmExp, index = 11, type = 'overlay')
```

```{r, plotSRM_facet, fig.width = 6, fig.height = 4,  fig.align = 'center'}
plotSRM(srmExp, index = 11, type = 'facet')

```


#### plotParent 

`plotParent` use a numeric value of parent mass as an input, and plots all SRM transitions which feature the selected parent mass. 

```{r, plotParent,fig.width = 6, fig.height = 4,  fig.align = 'center'}
plotParent(srmExp, 187.026)
```


### Peak Detection

`sRm` utilises the `xcms::peaksWithCentWave()` function to detection chromatogram features. 

```{r, peak_detection}
srmExpPeaks <-
  detectPeaks(srmExp,
              snthresh = 2,
              peakwidth = c(5,60))
```


The `plotPeakArea` can then be used to determine the quality of the peak integration. 

```{r, plotArea, fig.width = 6, fig.height = 4,  fig.align = 'center'}
plotPeakArea(srmExpPeaks, index = 11, sampleName = 'QC02')
```


### Peak Measures






